Memory-aware Adaptive Scheduling of Scientific Workflows on Heterogeneous Architectures\\

### 2025-05-16 main.tex
Done on my side
corrected formula of Res and description, committing also the IEEE footnote

\newcommand{\new}[1]{{\color{blue}#1}}

@@

@@

\section{Introduction} %: \skug{Full: 0, Polished: 0}}

%%% CONTEXT %%%
The analysis of massive datasets, originating from fields such as genomics, 
remote sensing, or biomedical imaging -- to name just a few -- has become ubiquitous in science;
this often takes the form of workflows, \iec separate software components chained together
in some kind of complex pipeline~\cite{DBLP:journals/dbsk/LeserHDEGHKKKKK21}.
These workflows are usually represented as directed acyclic graphs (DAGs).
The DAG vertices represent the software components (or, more generally, the workflow \emph{tasks}),
while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liu2018survey}.
Large workflows with resource-intensive tasks can easily exceed the capabilities of a 
single computer and are therefore executed on a parallel or distributed platform.
An efficient execution of the workflows on such platforms requires the mapping of tasks
to specific processors.
\new{Moreover, a task schedule -- \iec a valid execution order that respects the dependencies --
and possibly also starting times for the tasks are needed to increase utilization by reusing 
processors that have become idle after finishing a task.}

@@

\smallskip
\noindent{\bf Compute task ranks. }

@@

\smallskip
\noindent  {\bf Task assignment. }

@@

\smallskip
\noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
$$Res = availM_j - m_v - \!\!\!\! \sum_{u \in \Pi(v), u\notin T(p_j)}  \!\!\!\!\!\!\!\!\{c_{u,v}\}
- \sum_{w\in Succ(v)} \!\!\!\!\!\! \{c_{v,w}\}.$$

$T(p_j)$ is the set of tasks already scheduled on $p_j$; by Step 1, their files are
already in the memory of~$p_j$. However, the files from the
other predecessor tasks must be loaded into memory before executing task~$v$,
as well as the task data of size $m_v$, and the data generated for all successor tasks.
$Res$ then checks whether there is enough memory; if it is negative,
we have exceeded the memory of~$p_j$ with this tentative assignment.
%
In this case ($Res <0$), we try to evict some data from memory so that there is enough memory to execute task~$v$.
Clearly, we need to evict at least $Res$ data.
To this end, we propose a greedy approach, which evicts the largest files of $\PD_j$ until data of size $Res$ have been evicted. 
A variant where the smallest files are evicted first has been tested; it led to comparable results. 
%    
%    in order to avoid costly communications.
%    \AB{FYI We initially discussed evicting the largest files, but this leads to
%    large communications and does not seem efficient after all... Maybe we can think of another
%    approach that would take into account both data size and bottom level...}
While tentatively evicting files, we remove them from the list of pending memories and move them into a list
of memories pending in the communication buffer.
We keep track of the available buffer size, too -- every time a file is moved into the pending buffer, 
the available buffer size is reduced by the file size.

@@

\section{Dynamic scenario}
\label{sec:dyn}
%
In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
This information may include memory usage, running time, graph structures, or the status of the underlying infrastructure.
In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
By incorporating dynamic monitoring values, e.g., the resources a task consumed, the runtime environment can incorporate the data into the prediction model to provide more accurate resource predictions.
Also the underlying infrastructure can change during the workflow execution.
Examples are processor failures, node recoveries, or acquisition of new nodes.
Even if the hardware infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during workflow runtime, the previous schedule may become invalid, so that a new one must be calculated.

@@

\subsubsection{Workflow instances}
We run experiments on a set of several real-world workflows from Ref.~\cite{lotaru}
(atacseq, bacass, chipseq, eager, and methylseq). We use the WFGen generator~\cite{COLEMAN202216} to
create larger variants of the workflows above and add the instances created this way to the set.

@@

\subsubsection{Scheduling on the default cluster}
%
%    \paragraph{Normal cluster}
%
On the default cluster,  the three memory-aware heuristics are able to schedule all workflows
(see Figure~\ref{fig:success-rates-large}), while
the baseline \heft has a success rate of only $24.2\%$ ($75.7\%$ failure rate).
Indeed, \heft is only able to schedule small workflows; it cannot schedule any workflow 
with more than $4000$ tasks correctly as some tasks run out of memory. As soon as we are not in a setting
with abundant processing resources for small workflows, it is hence necessary to adopt
a memory-aware strategy in order to produce valid schedules.
%    Small workflows are less constrained by the memory requirement, so in the situation of abundant processing resources, even
%    a not memory-aware approach is able to produce valid schedules.
%    However, this advantage quickly vanishes as workflows grow larger.

@@

\subsubsection{Scheduling on the memory-constrained cluster}
\label{ss:mem-constrained-cluster}
%
On the memory-constrained cluster, \heft produces valid assignments in only $14$ out of $290$ experiments ($4.8\%$ success rate).
The successful schedules are achieved exclusively on the tiny workflows (with only two 200-task size-increased
workflow among them).
\heftbl successfully schedules $38\%$ of the workflows, \heftblc is able to schedule $49\%$ of them, while \heftmm
still schedules all of them, including even the largest ones, see Figure~\ref{fig:success-rates-tiny}.
As also observed on the default cluster, \heftmm seems to be less affected by the workflow size
than the other heuristics.

@@

\begin{figure}[tb]
    \centering
    \vspace{-0.3cm}
    \includegraphics[width=1\columnwidth] {images/ms_relations_by_wf_size-constrained-barplot}
    \caption{Relative makespans on the memory-constrained cluster.
%        produced by the heuristics to the makespan produced by the baseline, by workflow size.
    Smaller is better.}
    \label{fig:ms-relations-by-workflow-constrained}
\end{figure}

@@

\begin{figure}[tb]
    \centering
    \includegraphics[width=1\columnwidth] {images/runtimes-logarithmic2}
    \caption{Average running times (in seconds) of the heuristics \wrt workflow size. The $y$-axis is logarithmic.}
    \label{fig:runtimes-log}
%    \vspace{-0.3cm}
\end{figure}

>>>>

%\newcommand{\new}[1]{{\color{blue}#1}}
\newcommand{\new}[1]{{#1}}

@@

 \IEEEpubid{\begin{minipage}{\textwidth}\ \\[12pt]
 © 2025 IEEE. Personal use is permitted, but republication/redistribution requires IEEE permission.
 See https://www.ieee.org/publications/rights/index.html for more information.
 The published conference version of this paper appears in the Proceedings of 25th International Symposium on Cluster, Cloud and Internet Computing (CCGrid), IEEE Computer Society.
\end{minipage}}

@@

\section{Introduction} %: \skug{Full: 0, Polished: 0}}
    \IEEEpubidadjcol
%%% CONTEXT %%%
The analysis of massive datasets, originating from fields such as genomics, 
remote sensing, or biomedical imaging -- to name just a few -- has become ubiquitous in science;
this often takes the form of workflows, \iec separate software components chained together
in some kind of complex pipeline~\cite{DBLP:journals/dbsk/LeserHDEGHKKKKK21}.
These workflows are usually represented as directed acyclic graphs (DAGs).
The DAG vertices represent the software components (or, more generally, the workflow \emph{tasks}),
while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liu2018survey}.
Large workflows with resource-intensive tasks can easily exceed the capabilities of a 
single computer and are therefore executed on a parallel or distributed platform.
An efficient execution of the workflows on such platforms requires the mapping of tasks
to specific processors.
\new{Moreover, a task schedule -- \iec a valid execution order that respects the dependencies --
and possibly also starting times for the tasks are needed to increase utilization by reusing 
processors that have become idle after finishing a task.}
    \IEEEpubidadjcol

@@

\medskip
\noindent{\bf Compute task ranks. }

@@

\bigskip
\noindent  {\bf Task assignment. }

@@

\smallskip
\noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
$$Res = availM_j - \max{ \{  m_v,  \!\!\!\! \sum_{u \in \Pi(v), u\notin T(p_j)}  \!\!\!\!\!\!\!\!\{c_{u,v}\}
, \sum_{w\in Succ(v)} \!\!\!\!\!\! \{c_{v,w}\} \} }.$$

$T(p_j)$ is the set of tasks already scheduled on $p_j$; by Step 1, their files are
already in the memory of~$p_j$. However, the files from the
other predecessor tasks must be loaded into memory before executing task~$v$.
Depending on what part of $r_u$ was the largest, we need to have enough memory to store
the task data of size $m_v$, or the data generated for all successor tasks.
$Res$ then checks whether there is enough memory; if it is negative,
we have exceeded the memory of~$p_j$ with this tentative assignment.
%
In this case ($Res <0$), we try to evict some data from memory so that there is enough memory to execute task~$v$.
Clearly, we need to evict at least $Res$ data.
To this end, we propose a greedy approach, which evicts the largest files of $\PD_j$ until data of size $Res$ have been evicted. 
A variant where the smallest files are evicted first has been tested; it led to comparable results. 
%    
%    in order to avoid costly communications.
%    \AB{FYI We initially discussed evicting the largest files, but this leads to
%    large communications and does not seem efficient after all... Maybe we can think of another
%    approach that would take into account both data size and bottom level...}
While tentatively evicting files, we remove them from the list of pending memories and move them into a list
of memories pending in the communication buffer.
We keep track of the available buffer size, too -- every time a file is moved into the pending buffer, 
the available buffer size is reduced by the file size.

@@

\vspace{-0.09cm}
\section{Dynamic scenario}
\label{sec:dyn}
%
In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
This information may include memory usage, running time, graph structures, or the status of the underlying infrastructure.
In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
By incorporating dynamic monitoring values, e.g., the resources a task consumed, the runtime environment can incorporate the data into the prediction model to provide more accurate resource predictions.
Also the underlying infrastructure can change during the workflow execution.
Examples are processor failures, node recoveries, or acquisition of new nodes.
Even if the hardware infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during workflow runtime, the previous schedule may become invalid, so that a new one must be calculated.

@@

\medskip
\subsubsection{Workflow instances}
We run experiments on a set of several real-world workflows from Ref.~\cite{lotaru}
(atacseq, bacass, chipseq, eager, and methylseq). We use the WFGen generator~\cite{COLEMAN202216} to
create larger variants of the workflows above and add the instances created this way to the set.

@@

\medskip
\subsubsection{Scheduling on the default cluster}
%
%    \paragraph{Normal cluster}
%
On the default cluster,  the three memory-aware heuristics are able to schedule all workflows
(see Figure~\ref{fig:success-rates-large}), while
the baseline \heft has a success rate of only $24.2\%$ ($75.7\%$ failure rate).
Indeed, \heft is only able to schedule small workflows; it cannot schedule any workflow 
with more than $4000$ tasks correctly as some tasks run out of memory. As soon as we are not in a setting
with abundant processing resources for small workflows, it is hence necessary to adopt
a memory-aware strategy in order to produce valid schedules.
%    Small workflows are less constrained by the memory requirement, so in the situation of abundant processing resources, even
%    a not memory-aware approach is able to produce valid schedules.
%    However, this advantage quickly vanishes as workflows grow larger.

@@

\medskip
\subsubsection{Scheduling on the memory-constrained cluster}
\label{ss:mem-constrained-cluster}
%
On the memory-constrained cluster, \heft produces valid assignments in only $14$ out of $290$ experiments ($4.8\%$ success rate).
The successful schedules are achieved exclusively on the tiny workflows (with only two 200-task size-increased
workflow among them).
\heftbl successfully schedules $38\%$ of the workflows, \heftblc is able to schedule $49\%$ of them, while \heftmm
still schedules all of them, including even the largest ones, see Figure~\ref{fig:success-rates-tiny}.
As also observed on the default cluster, \heftmm seems to be less affected by the workflow size
than the other heuristics.

@@

\begin{figure}[tb]
    \centering
    \vspace{-0.3cm}
    \includegraphics[width=1.03\columnwidth] {images/ms_relations_by_wf_size-constrained-barplot}
    \caption{Relative makespans on the memory-constrained cluster.
%        produced by the heuristics to the makespan produced by the baseline, by workflow size.
    Smaller is better.}
    \label{fig:ms-relations-by-workflow-constrained}
\end{figure}

@@

\begin{figure}[tb]
    \centering
  \hspace{-.5cm}  \includegraphics[width=1.01\columnwidth] {images/runtimes-logarithmic2}
    \caption{Average running times (in seconds) of the heuristics \wrt workflow size. The $y$-axis is logarithmic.}
    \label{fig:runtimes-log}
%    \vspace{-0.3cm}
\end{figure}
